Urban and socio-economic correlates of property prices in Dublin's area

### 2020-08-18 manuscript/manuscript.Rmd
paper submission

# libraries --------------------------------------------------------------------
library(rticles)
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)
library(papaja)
library(patchwork)
library(kableExtra)
library(latex2exp)
library(yardstick)
library(sf)
library(tmap)
library(lwgeom)
library(osmdata)
library(OpenStreetMap)
library(mgcv)
library(xgboost)

@@

# data -------------------------------------------------------------------------
geocode_dublin_2018 <- here::here("data/ppr_dublin_2018.csv") %>% 
  readr::read_csv()

house_dublin_2018 <- geocode_dublin_2018 %>% 
  dplyr::filter(!grepl('apartment|Apartment|APART|APT|Apt', Address))

final_dublin_2018 <- here::here("data/data_dublin_2018_osm_features.rds") %>% 
  readr::read_rds()

# data path --------------------------------------------------------------------
data_path <- here::here() %>% 
  dirname() %>% 
  dirname() %>% 
  paste0("/data/property_price_ireland_data")

# all price registry------------------------------------------------------------
data_dublin_geocoded <- file.path(data_path, "data_dublin_geocoded_clean.rds") %>% 
  readr::read_rds() %>%
  janitor::clean_names() %>%
  dplyr::mutate(
    year_fact = as.factor(as.character(year))
  ) %>% 
  dplyr::filter(year == 2018)

mean_dat <- mean(data_dublin_geocoded$price)
sd_dat <- sd(data_dublin_geocoded$price)

data_dublin_geocoded <- data_dublin_geocoded %>%
  dplyr::filter(price > mean_dat - sd_dat)

# all dublin features ----------------------------------------------------------
data_dublin_2018_features <- data_path %>%
  file.path("data_dublin_2018_osm_features.rds") %>%
  readr::read_rds() %>% 
  janitor::remove_empty(which = "cols") %>% 
  janitor::clean_names() %>%
  dplyr::mutate(geog_id = stringr::str_remove(geogid, "[A]"))

list_geogid <- unique(data_dublin_2018_features$geog_id)

data_dublin_2018_features <- c(
  "carers.csv",
  "disabilty_age_group.csv", 
  "general_health.csv",
  "housing_occupancy.csv",
  "housing_rooms.csv",
  "housing_tenure.csv",
  "housing_type.csv",
  "population.csv",
  "religion.csv"
  ) %>%
  purrr::map(read_join) %>% 
  purrr::reduce(dplyr::left_join) %>% 
  dplyr::inner_join(data_dublin_2018_features, by = c("GEOG_ID" = "geog_id"))

# train test split -------------------------------------------------------------
train_index <- sample(1:nrow(data_dublin_2018_features), 0.8 * nrow(data_dublin_2018_features))
test_index <- setdiff(1:nrow(data_dublin_2018_features), train_index)

@@

Since 2010, under the Irish Property Services (Regulation) Act, all individuals acquiring a property in Ireland have to declare it to Property Services Regulatory Authority (PSRA). They must provide details such as the date of sale, the price and the address of all residential properties purchased in Ireland as declared to the Revenue Commissioners for stamp duty purposes [@rppr2020]. Data is filed electronically by persons doing the conveyancing of the property on behalf of the purchaser, and it must be noted that errors may occur when the data is being filed.

Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only properties sold in 2018 were processed to avoid potential temporal incoherence. The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is `r nrow(geocode_dublin_2018)`. From these, all property corresponding to apartments were removed (`r scales::percent(1-(nrow(house_dublin_2018)/nrow(geocode_dublin_2018)), accuracy = 0.1)` of the original dataset). Finally, unsuccessfully geocoded properties were also removed from the original dataset (`r scales::percent(1-(nrow(final_dublin_2018)/nrow(house_dublin_2018)), accuracy = 0.1)`) to obtain the `r nrow(final_dublin_2018)` of the final dataset used to identify the urban and socio-economic correlates of property prices.

@@

@@

# knots ------------------------------------------------------------------------
# boundaries_dublin <- 
#   osmdata::opq(bbox = 'Dublin, Ireland') %>%
#   osmdata::add_osm_feature(key = 'admin_level', value = '7') %>%
#   osmdata::osmdata_sf() %>% 
#   osmdata::unique_osmdata()
# 
# bb <- st_bbox(boundaries_dublin$osm_multipolygons$geometry)
# 
# readr::write_rds(bb, here::here("data/bb.rds"))

bb <- readr::read_rds(here::here("data/bb.rds"))

@@

# model ------------------------------------------------------------------------
gam_dublin_gps <- mgcv::gam(
  price ~ s(lng, lat, bs = "so", xt = list(bnd = bound)), 
  data = dplyr::filter(data_dublin_geocoded, year == 2018), 
  method = "REML", 
  knots = knots,
  family = gaussian(link = "identity")
)

gam_pred <- gam_density(gam_dublin_gps, too.far = 0.05, n.grid = 30)
```

In 2018, the average price of a sold property in Dublin was `r scales::dollar(round(mean_dat, 0), suffix = "", prefix = "€")` with a standard deviation of `r scales::dollar(round(sd_dat, 0), suffix = "", prefix = "€")`. In order to remove potential human errors and outliers, prices higher or lower than 1 SD were removed from the original dataset.

```{r distrib-plot, fig.height=3, out.width="0.98\\columnwidth", fig.pos ="!h", fig.cap="(ref:distrib-plot)"}
data_dublin_geocoded %>%
  dplyr::filter(year == 2018) %>%
  magrittr::use_series(price) %>%
  stats::density() %$%
  data.frame(x, y) %>%
  ggplot() +
  geom_segment(aes(
    x = x,
    xend = x,
    y = 0,
    yend = y,
    color = x
  )) +
  scale_x_continuous("Price", labels = scales::dollar_format(suffix = "", prefix = "€")) +
  scale_y_continuous("Density") +
  scale_color_gradient(
    low = '#3a53a4', 
    high = '#ed1c24', 
    ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 12),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )
```

```{r distrib-gps, fig.height=5, out.width="0.98\\columnwidth", fig.pos ="!h", fig.cap="(ref:distrib-gps)"}
gam_pred %>%
  dplyr::filter(value > 200000) %>%
  ggplot() +
  geom_raster(aes(x = lng, y = lat, fill = value)) +
  geom_sf(data = dublin_sea, fill = "light blue") +
  scale_fill_gradient(
    "Price Density", 
    low = '#3a53a4', 
    high = '#ed1c24', 
    trans = "log",
    breaks = c(250000, 1000000), 
    labels = scales::dollar_format(suffix = "", prefix = "€")) +
  scale_x_continuous(
    "Longitude", 
    limits = c(-6.5, -6), 
    breaks = seq(-6.5, -6, 0.2)) +
  scale_y_continuous(
    "Latitude", 
    limits = c(53.2, 53.5), 
    breaks = seq(53.2, 53.5, 0.1)) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 12),
    legend.position = "none"
  )
```

The density of housing prices reveals a unimodal distribution of property prices (Figure \@ref(fig:distrib-plot)). In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \@ref(fig:distrib-gps)).

@@

metrics_result <- result %>% 
  yardstick::metrics(price, pred) %>% 
  select(-.estimator) %>% 
  mutate(.estimate = round(.estimate, 2))

@@

Using the XGBoost algorithm with a tree based booster and a tweedie regression with log-link learning objective, the 160 urban features are explaining `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.

@@

```{r osm-features-xgb, fig.cap="(ref:osm-features-xgb)", fig.height=7, fig.pos ="!h", out.width="0.98\\columnwidth"}
result_accuracy <- result %>%
  ggplot(aes(x = price, y = pred)) +
  geom_point(alpha = 0.1) +
  geom_smooth(
    method = "lm",
    fullrange = TRUE,
    se = TRUE,
    color = "black",
    size = 0.1
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 12),
    legend.position = "none"
  ) +
  labs(
    title = "A",
    subtitle = latex2exp::TeX(paste(
      "$R^2 =$", 
      scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"]),
      ";",
      "$RMSE =$",
      scales::dollar(
        metrics_result$.estimate[metrics_result$.metric == "rmse"], 
        suffix = "", 
        prefix = "€"
      ),
      ";",
      "$MAE =$",
      scales::dollar(
        metrics_result$.estimate[metrics_result$.metric == "mae"], 
        suffix = "", 
        prefix = "€"
      )
      )
    )
  )

result_error <- result %>%
  dplyr::mutate(diff_percent = (pred - price) / price) %>%
  ggplot(aes(x = price, y = diff_percent)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(
    color = "black",
    se = TRUE,
    fullrange = TRUE,
    size = 0.1
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price Error",
    limits = c(-0.5, 0.5),
    labels = scales::percent
  ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 12)
  ) +
  ggtitle("B")

@@

```{r osm-features-table, results="asis"}
table_osm %>% 
  knitr::kable(
    "latex",
    caption = "(ref:osm-features-table)",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kableExtra::kable_styling(
    font_size = 8, 
    full_width = TRUE, 
    latex_options = "HOLD_position"
  )
```

@@

comparison_xgboost %>% 
  dplyr::select(-booster) %>% 
  knitr::kable(
    "latex",
    caption = "(ref:census-model-comparison)",
    booktabs = T,
    linesep = "",
    col.names = c("Objective", "$RMSE$", "$R^2$", "$MAE$"),
    escape = FALSE
  ) %>%
  kableExtra::kable_styling(font_size = 8, latex_options = "HOLD_position") %>% 
  pack_rows(index = c("Tree based" = 4, "Linear regression based" = 4)) %>% 
  kableExtra::footnote("Missing value ares due to incompatibilities between data and methods.")
```

@@

```{r census-features-xgb, fig.cap="(ref:census-features-xgb)", fig.height=7, fig.pos ="!h", out.width="0.98\\columnwidth"}
result_accuracy <- result %>%
  ggplot(aes(x = price, y = pred)) +
  geom_point(alpha = 0.1) +
  geom_smooth(
    method = "lm",
    fullrange = TRUE,
    se = TRUE,
    color = "black",
    size = 0.1
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 12),
    legend.position = "none"
  ) +
    labs(
    title = "A",
    subtitle = latex2exp::TeX(paste(
      "$R^2 =$", 
      scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"]),
      ";",
      "$RMSE =$",
      scales::dollar(
        metrics_result$.estimate[metrics_result$.metric == "rmse"], 
        suffix = "", 
        prefix = "€"
      ),
      ";",
      "$MAE =$",
      scales::dollar(
        metrics_result$.estimate[metrics_result$.metric == "mae"], 
        suffix = "", 
        prefix = "€"
      )
      )
    )
  )

result_error <- result %>%
  dplyr::mutate(diff_percent = (pred - price) / price) %>%
  ggplot(aes(x = price, y = diff_percent)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(
    color = "black",
    se = TRUE,
    fullrange = TRUE,
    size = 0.1
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price Error",
    limits = c(-0.5, 0.5),
    labels = scales::percent
  ) +
  theme_apa() +
  theme(text = element_text(family = "serif", size = 12)) +
  ggtitle("B")

@@

Using the XGBoost algorithm with a tree based booster and a gamma regression with log-link learning objective, the 48 socio-economic features explain `r scales::percent(metrics_result$.estimate[metrics_result$.metric == "rsq"])` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "rmse"], suffix = "", prefix = "€")` and a $MAE$ of `r scales::dollar(metrics_result$.estimate[metrics_result$.metric == "mae"], suffix = "", prefix = "€")`.

@@

```{r census-features-table, results="asis"}
table_census %>%
  knitr::kable(
    "latex",
    caption = "(ref:census-features-table)",
    booktabs = T,
    linesep = ""
  ) %>%
  kableExtra::kable_styling(
    font_size = 8, 
    full_width = TRUE, 
    latex_options = "HOLD_position"
  )
```

@@

While Dublin is a very specific city due to its low population density spread on a large surface, its housing market is one of the most expensive in Europe. The main reason is the structure of the market, mainly made of residential properties and few apartment buildings. In this study, the prices of `r nrow(data_dublin_2018_features)` properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority dataset in order to identify the potential factors correlating with these prices. The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market [@kitchin2013making]. However, a major limitation of this study is the absence of characteristics for the property itself. Indeed, prices are mainly determined by factors such as the size of the property or the number of rooms. In addition, while some features appear to be more important than others in their contribution to property prices, these results are not causal. Thus, features' importance can be either the cause or the consequence of property prices in a neighborhood.

>>>>

# libraries --------------------------------------------------------------------
library(rticles)
library(scales)
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)
library(papaja)
library(patchwork)
library(kableExtra)
library(latex2exp)
library(yardstick)
library(sf)
library(tmap)
library(lwgeom)
library(osmdata)
library(OpenStreetMap)
library(mgcv)
library(xgboost)

@@

# data -------------------------------------------------------------------------
n_raw <- here::here("data/ppr_dublin_2018.csv") %>% 
  readr::read_csv() %>% 
  nrow()

n_house <- here::here("data/ppr_dublin_2018.csv") %>% 
  readr::read_csv() %>% 
  dplyr::filter(!grepl('apartment|Apartment|APART|APT|Apt', Address)) %>% 
  nrow()

n_final <- here::here("data/data_dublin_2018_features.rds") %>% 
  readr::read_rds() %>% 
  nrow()

# all dublin features ----------------------------------------------------------
data_dublin_2018_features <- here::here("data/data_dublin_2018_features.rds") %>%
  readr::read_rds()

# train test split -------------------------------------------------------------
train_index <- sample(1:n_final, 0.8 * n_final)
test_index <- setdiff(1:n_final, train_index)

@@

Since 2010, under the Irish Property Services (Regulation) Act, all individuals acquiring a property in Ireland have to declare it to Property Services Regulatory Authority (PSRA). They must provide details such as the date of sale, the price and the address of all residential properties purchased in Ireland as declared to the Revenue Commissioners for stamp duty purposes [@rppr2020]. Data is filed electronically by persons doing the conveyancing of the property on behalf of the purchaser, and it must be noted that errors may occur when the data is being filed. 
Because of the evolution of urban planning and the evolution of socio-economic features measured by the 2016 census, only properties sold in 2018 were processed to avoid potential temporal incoherence. The total amount of property sold in 2018 in the county Dublin which appears in the Property Price Register is `r n_raw`. From these, all property corresponding to apartments were removed (`r scales::percent(1-(n_house/n_raw), accuracy = 0.1)` of the original dataset). Finally, unsuccessfully geocoded properties were also removed from the original dataset (`r scales::percent(1-(n_final/n_house), accuracy = 0.1)`) to obtain the `r n_final` of the final dataset used to identify the urban and socio-economic correlates of property prices.

@@

```{r descriptives}
data_mean <- data_dublin_2018_features$price %>% 
  mean(na.rm = TRUE) %>% round(0) %>% scales::dollar(suffix = "", prefix = "€")
data_sd <- data_dublin_2018_features$price %>% 
  sd(na.rm = TRUE) %>% round(0) %>% scales::dollar(suffix = "", prefix = "€")
```

In 2018, the average price of a sold property in Dublin was `r data_mean` with a standard deviation of `r data_sd` (Figure \@ref(fig:distrib-plot)).

@@

# knots ------------------------------------------------------------------------
bb <- readr::read_rds(here::here("data/bb.rds"))

@@

# model ------------------------------------------------------------------------
gam_dublin_gps <- mgcv::gam(
  price ~ s(lng, lat, bs = "so", xt = list(bnd = bound)), 
  data = data_dublin_2018_features, 
  method = "REML", 
  knots = knots,
  family = gaussian(link = "identity")
)

gam_pred <- gam_density(gam_dublin_gps, too.far = 0.05, n.grid = 100)
```

```{r distrib-plot, fig.height=4, out.width="0.98\\columnwidth", fig.pos ="!h", fig.cap="(ref:distrib-plot)"}
data_dublin_2018_features %>%
  magrittr::use_series(price) %>%
  stats::density() %$%
  data.frame(x, y) %>%
  ggplot() +
  geom_segment(aes(
    x = x,
    xend = x,
    y = 0,
    yend = y,
    color = x
  )) +
  scale_x_continuous("Price", labels = dollar_format(suffix = "", prefix = "€")) +
  scale_y_continuous("Density") +
  scale_color_gradient(
    low = '#3a53a4', 
    high = '#ed1c24', 
    ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 12),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )
```

```{r distrib-gps, fig.height=5, out.width="0.98\\columnwidth", fig.pos ="!h", fig.cap="(ref:distrib-gps)"}
gam_pred %>%
  dplyr::filter(value > 200000) %>%
  ggplot(aes(x = lng, y = lat)) + 
  geom_raster(aes(fill = value), na.rm = TRUE) +
  geom_contour(aes(z = value), color = "black", na.rm = TRUE) +
  geom_sf(data = dublin_sea, fill = "light blue", inherit.aes = FALSE) +
  scale_fill_gradient(
    "Price Density", 
    low = '#3a53a4', 
    high = '#ed1c24', 
    trans = "log",
    breaks = c(250000, 1000000), 
    labels = scales::dollar_format(suffix = "", prefix = "€"))  +
  scale_x_continuous(
    "Longitude", 
    limits = c(-6.5, -6), 
    breaks = seq(-6.5, -6, 0.2)) +
  scale_y_continuous(
    "Latitude", 
    limits = c(53.2, 53.5), 
    breaks = seq(53.2, 53.5, 0.1)) +
  theme_classic() +
  theme(
    text = element_text(family = "serif", size = 12),
    legend.position = "none",
    panel.background = element_rect(fill = "grey50")
  )
```

The density of housing prices reveals a unimodal distribution of property prices. In addition, the geographical distribution of property prices indicates a higher valuation in the South-West of the city as well as on the coast line (Figure \@ref(fig:distrib-gps)).

@@

metrics_result <- result %>% 
  yardstick::metrics(price, pred) %>% 
  select(-.estimator) %>% 
  mutate(.estimate = case_when(
    .metric == "rsq" ~ scales::percent(.estimate, accuracy = 0.1),
    TRUE ~ scales::dollar(.estimate, suffix = "", prefix = "€")
  )) %>% 
  dplyr::pull(.estimate, .metric)

@@

Using the XGBoost algorithm with a tree based booster and a tweedie regression with log-link learning objective, the 160 urban features are explaining `r metrics_result["rsq"]` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r metrics_result["rmse"]` and a $MAE$ of `r metrics_result["mae"]`.

@@

```{r osm-features-xgb, fig.cap="(ref:osm-features-xgb)", fig.height=7, fig.pos ="!h", out.width="0.98\\columnwidth"}
result_accuracy <- result %>%
  ggplot(aes(x = price, y = pred)) +
  geom_point(alpha = 0.1) +
  geom_smooth(
    method = "lm",
    fullrange = TRUE,
    se = TRUE,
    color = "black",
    size = 0.1,
    na.rm = TRUE
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 12)
  ) +
  ggtitle("A")

result_error <- result %>%
  dplyr::mutate(diff_percent = (pred - price) / price) %>%
  ggplot(aes(x = price, y = diff_percent)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(
    color = "black",
    se = TRUE,
    fullrange = TRUE,
    size = 0.1,
    na.rm = TRUE
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price Error",
    limits = c(-0.5, 0.5),
    labels = scales::percent
  ) +
  theme_apa() +
  theme(
    text = element_text(family = "serif", size = 12)
  ) +
  ggtitle("B")

@@

```{r osm-features-table, results="asis"}
table_osm %>% 
  knitr::kable(
    "latex",
    caption = "(ref:osm-features-table)",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kableExtra::kable_styling(
    font_size = 8, 
    full_width = TRUE, 
    latex_options = "HOLD_position"
  ) %>% 
  kableExtra::column_spec(2, width = "10em")
```

@@

comparison_xgboost %>% 
  dplyr::select(-booster) %>% 
  knitr::kable(
    "latex",
    caption = "(ref:census-model-comparison)",
    booktabs = T,
    linesep = "",
    col.names = c("Objective", "$RMSE$", "$R^2$", "$MAE$"),
    escape = FALSE
  ) %>%
  kableExtra::kable_styling(font_size = 8, latex_options = "HOLD_position") %>% 
  kableExtra::pack_rows(index = c("Tree based" = 4, "Linear regression based" = 4)) %>% 
  kableExtra::footnote("Missing value ares due to incompatibilities between data and methods.")
```

@@

```{r census-features-xgb, fig.cap="(ref:census-features-xgb)", fig.height=7, fig.pos ="!h", out.width="0.98\\columnwidth"}
result_accuracy <- result %>%
  ggplot(aes(x = price, y = pred)) +
  geom_point(alpha = 0.1) +
  geom_smooth(
    method = "lm",
    fullrange = TRUE,
    se = TRUE,
    color = "black",
    size = 0.1,
    na.rm = TRUE
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  theme_apa() +
  theme(text = element_text(family = "serif", size = 12)) +
  ggtitle("A")

result_error <- result %>%
  dplyr::mutate(diff_percent = (pred - price) / price) %>%
  ggplot(aes(x = price, y = diff_percent)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(
    color = "black",
    se = TRUE,
    fullrange = TRUE,
    size = 0.1,
    na.rm = TRUE
  ) +
  scale_x_continuous(
    "Property Price",
    labels = scales::dollar_format(suffix = "", prefix = "€")
  ) +
  scale_y_continuous(
    "Predicted Property Price Error",
    limits = c(-0.5, 0.5),
    labels = scales::percent
  ) +
  theme_apa() +
  theme(text = element_text(family = "serif", size = 12)) +
  ggtitle("B")

@@

Using the XGBoost algorithm with a tree based booster and a gamma regression with log-link learning objective, the 48 socio-economic features explain `r metrics_result["rsq"]` of the property price variance (`r lm_result$statistic$modelfit$r2`) with a $RMSE$ of `r metrics_result["rmse"]` and a $MAE$ of `r metrics_result["mae"]`.

@@

```{r census-features-table, results="asis"}
table_census %>%
  knitr::kable(
    "latex",
    caption = "(ref:census-features-table)",
    booktabs = T,
    linesep = ""
  ) %>%
  kableExtra::kable_styling(
    font_size = 8, 
    full_width = TRUE, 
    latex_options = "HOLD_position"
  ) %>% 
  kableExtra::column_spec(2, width = "10em") 
```

@@

While Dublin is a very specific city due to its low population density spread on a large surface, its housing market is one of the most expensive in Europe. The main reason is the structure of the market, mainly made of residential properties and few apartment buildings. In this study, the prices of `r n_final` properties sold in 2018 in the area of Dublin were analysed from the Property Services Regulatory Authority dataset in order to identify the potential factors correlating with these prices. The enforcement of a housing price registry has been shown as being an essential tool to regulate the housing market [@kitchin2013making]. However, a major limitation of this study is the absence of characteristics for the property itself. Indeed, prices are mainly determined by factors such as the size of the property or the number of rooms. In addition, while some features appear to be more important than others in their contribution to property prices, these results are not causal. Thus, features' importance can be either the cause or the consequence of property prices in a neighborhood.
